CDSR Human Data/Modelling Paper

### 2013-09-12 scenarios/cdsr/trunk/papers/ACS2013/main.tex
tables 1 and 2 together

\documentclass[11pt,letterpaper]{article}
\usepackage{cogsys}
\usepackage{cogsysapa}
% \usepackage{apacite}
% \usepackage{graphicx}
\usepackage[T1]{fontenc}
\usepackage{times}
\usepackage[pdftex]{graphicx} % use this when importing PDF files
\usepackage{caption}
\usepackage{subcaption}
\usepackage{multirow}

@@

\subsection{Research Questions, Design and Stimuli}

@@

Both of these research questions address the fundamental claim that CDSRs are contextually defined. Question 1 addresses this claim by examining variation in CDSR placement within different situtations. In a context exhibiting clear contextual clues (e.g. determining a safe zone in a battle context where all information is known and battle formations are well-defined) one would expect high agreement in human identification of the CDSR. In a less well-defined context, one might expect little to no agreement between human communicators.  Research question 2 addresses the contextual nature of CDSRs in terms of whether or not the variation in the location and extent of a CDSR systematically follow the variations  in the given context.

@@

\subsection{Task 1 - Results}
We use two methods to analyze the results of region drawing task.  First, we use interannotator agreement for two reasons: (1) if there is fair to good overall agreement, this is evidence that people change the regions they produce based systematically with context, (2) we expect less agreement in ambiguous contexts (e.g., the circle (a) and U-shaped (c) of desk arrangements as well as the turret + 1 (l) military engagement configuration).  Second, we use the extent points of the regions to assess changes between trials.  That is, as the desks are moved back from stimuli (e) to (f), we expect the maximum x-coordinate of the region to increase.

@@

Figure \ref{fig:subj-kappa} shows a histogram of the per subject average kappa scores using ten equal width intervals covering the $\kappa$ range between 0 and 1.  It is evident from this figure that most subjects had a $\kappa$ score between $0.2-0.5$ with a overall majority of subjects having a $\kappa$ score between $0.3-0.4$. A $\kappa$ score between $0.3-0.4$ would be considered as indicating poor to fair agreement, i.e. there was some agreement between subjects but less that would be expected. However, this calculation was done across all the stimuli and as we have already mentioned some of the stimuli were deliberately designed to be non-prototypical, in particular we expected the classroom stimuli where the chairs were arranged in a circle (Figure \ref{fig:classroom-a}) and in a U shaped pattern (Figure \ref{fig:classroom-c} and its horizontally flipped equivalent) to be relatively novel contexts for the subjects. Indeed, if we look at the histogram of the average pairwise $\kappa$ scores by stimuli, Figure \ref{fig:stim-kappa}, it is evident that there is a large variation, ranging from $0$ to $0.8$, in $\kappa$ scores for each stimuli.

\begin{table}
\centerline{
\begin{tabular}{l c c}
Stim ID & Average Pairwise $\kappa$ & No Region Responses\\
\hline
\hline
(h) & 0.5034 & 0\\ % 1_2
(g) & 0.4585 & 1\\ % 1_1
(k) & 0.4548 & 0\\ % 3_3
(f) & 0.3930 & 0\\ % 2_4
(i) & 0.3906 & 0\\ % 3_1
(e-flipped) & 0.3813 & 2\\ % 2_3
(j) & 0.3787 & 1\\ % 3_2
(f-flipped) & 0.3720 & 2\\ % 2_5
(e) & 0.3629 & 1\\ %2_2
(l) & 0.3164& 0\\ % 3_4
(c) & 0.2438 & 2\\ % 2_6
(c-flipped) & 0.2414 & 4\\ %2_7
(a) & 0.0996 & 10 \\ % 2_1
\hline
\end{tabular}
}
\label{tab:stimuli-kappa-scores}
\caption{The average pairwise $\kappa$ scores and the number of no region response by visual stimuli. The stimuli id are based on the labels in Figure \ref{exp-stimuli} with the horizontally flipped versions of stimuli indicated.}
\end{table}

@@

In more detail, Table \ref{tab:stimuli-kappa-scores} lists the average pairwise $\kappa$ score for each of the visual stimuli used in the experiment and the number of no region responses for each stimuli. It is clear from this table that the non-prototypical stimuli have the lowest $\kappa$ scores and also that there is relationship between typicality and the number of no region responses. Given this relatively large variation across stimuli, which we attribute to the prototypicality of the depicted context, we computed the average pairwise $\kappa$ scores for each subject when the non-prototypical stimuli trials were excluded. Figure \ref{fig:subj-kappa-proto} presents the histogram of the resulting average pairwise subject  $\kappa$ scores by subject. Not surpringly, with the exclusion on the non-prototypical stimuli the average pairwise $\kappa$ scores increase, with over half the subjects having an average pairwise $\kappa$ score between $0.4$ and $0.6$. $\kappa$ scores in this range can be interpreted as fair to good.

@@

The analysis of the task 1 trials indicates that, over a range of domains and spatial regions, the responses from the subjects did agree, to a good degree. Furthermore, their results support the hypothesis that the defintion of CDSRs in a given context is effected by the location and orientation of objects in the environment. For example, focusing on the classroom scenes (stimuli, $a$, $c$, $e$, and $f$) the only difference between these trials was the layout of the chairs in the room and yet there is a substantial variation across these stimuli of the agreement between subjects on the specified spatial region, namely \textit{the front of the room}. Situation $f$ resulted in an average pairwise agreement of $0.393$ and in comparision context $a$ had an average pairwise agreement of $0.0996$.

@@

\begin{table}
\centerline{
\begin{tabular}{c c}
Expected Relationship & Human Results  \\
\hline
\hline
area(g) > area(h) & 291.2 > 276.2 (p<.003) \\
maximumX(f) > maximumX(e) & 176.1 > 167.1   (p<.47) \\
minimumX(e-flipped) > minimumX(f-flipped) & 336.1 > 321.3 (p<.12) \\
maximumY(i) > maxmimumY(j) & 304.64 > 265.04 (p<.03) \\
maximumY(j) > maxmimumY(k) & 265.04 > 251.71 (p<.47) \\
maximumY(k) = maxmimumY(l) & 265.04 != 290.03 (p<.04) \\
\hline
\end{tabular}
}
\label{tab:region-location-tests}
\caption{Comparison of the expected relationships to the mean human responses.}
\end{table}

@@

\section{Related Work}

As the results above show, in order for an artificial cognitive system to operate successfully with humans (e.g. understanding spatially-situated task utterances), they must have access to internal models which can both represent CDSRs and recognise them in new environments. Whilst spatial reasoning has been an active topic in AI and robotics for decades, CDSRs present a new problem which requires solutions beyond the current state-of-the-art. Regions are central to QSR work~\cite{Cohn:2001}. Regions can either have crisp boundaries, usually defined by geometric shapes (e.g. the bounding boxes of visually tracked objects~\cite{SridharCohn:10}), or vague boundaries created qualitatively by multiple crisp regions (e.g. the `egg-yolk model'~\cite{Cohn96b}), or by thresholding an underlying spatially-mapped function (e.g. those created by potential fields~\cite{brenneretal07ijcai} or by Gaussian kernels~\cite{burbridge-dearden12}). The fundamental decision when creating a computational model of a CDSR is the method used to represent its boundary. Given their context-dependent nature, boundary definitions based purely on single geometric shapes seem unlikely to model the regions with sufficient accuracy. However, arbitrary boundaries can be defined using a sequence of geometric primitives (usually straight lines), as in chain codes for image region representations~\cite{Freeman:1961}.

This is an approach we followed in previous work \cite{Hawes:2012}, where we developed an approach that defined CDSRs using \textit{anchor points} \cite{DBLP:journals/jetai/KlenkFTK11} as the vertices of the region polygon. Anchor points are symbolic descriptions that link a conceptual entity (e.g., the front of a classroom) to perceived entities (e.g., the lower left extent of a desk).  Once a particular representation has been modelled in this manner, we were able to find the same type of region in previously unseen rooms using analogical transfer. An example of the representation and transfer approach is Figure \ref{fig:anchor-point-transfer}.  In the know example, a human defines the front of the classroom as a polygon consisting of four anchor points.  Analogical transfer works by aligning structured representations and transferring partially aligned statements from the known example to the new environment.  In this case, the anchor points are transferred from the group of desks in the known example to the corresponding group of desks in the new environment.

\begin{figure}
  \includegraphics[width=\columnwidth]{figures/classroom-transfer.png}
  \caption{The known example defines the front of the room with four anchor points.  By transferring the anchor points to the new environment, the system is able to ground the room's front in its perception.}
  \label{fig:anchor-point-transfer}
\end{figure}

This method satisfies our criteria for a cognitive systems solution for the following reasons: (1) the alignment process integrates spatial and semantic knowledge about the environment, (2) analogy enables transfer from a single example, (3) while untested, by linking conceptual knowledge to perception, regions defined by anchor points should support any symbolic reasoning task. Applying this model to the points and regions indicated in our experiment illustrates two limitations.  Regions do not always physically bounded by objects.  In the classroom examples, people frequently did not extend the front of the room all way into the corners.  In the safety example, people did not indicate that it was safe behind the lone turret in front of the others.  This indicates that people consider the dynamics of the context when evaluating these terms, which is not accounted for in this model.

The use of spatial regions connected to function has been explored in other robotics work, e.g. ~\cite{Karg:2012,Fasola:2013}, but these regions are based on fields anchored at single points in the environment or on a particular object. Such fields can only vary with respect to context in terms of their size, which does not make them particularly well suited to modelling CDSRs.

>>>>

\documentclass[11pt,letterpaper]{article}
\usepackage{cogsys}
\usepackage{cogsysapa}
% \usepackage{apacite}
% \usepackage{graphicx}
\usepackage[T1]{fontenc}
\usepackage{times}
\usepackage[pdftex]{graphicx} % use this when importing PDF files
\usepackage{caption}
\usepackage{subcaption}
\usepackage{multirow}
% \usepackage{subfig}

@@

@@

Both of these research questions address the fundamental claim that CDSRs are contextually defined. Question 1 addresses this claim by examining variation in CDSR placement within different situations. In a context exhibiting clear contextual clues (e.g. determining a safe zone in a battle context where all information is known and battle formations are well-defined) one would expect high agreement in human identification of the CDSR. In a less well-defined context, one might expect little to no agreement between human communicators.  Research question 2 addresses the contextual nature of CDSRs in terms of whether or not the variation in the location and extent of a CDSR systematically follow the variations  in the given context.

@@

\subsection{Task 1 - Results}
We use two methods to analyze the results of region drawing task.  First, we use interannotator agreement for two reasons: (1) if there is fair to good overall agreement, this is evidence that people change the regions they produce based systematically with context, (2) we expect less agreement in ambiguous contexts (e.g., the circle (a) and U-shaped (c) desk arrangements as well as the turret + 1 (l) military engagement configuration).  Second, we use the extent points of the regions to assess changes between trials.  That is, as the desks are moved back from stimuli (e) to (f), we expect the maximum x-coordinate of the region to increase.

@@

%However, this calculation was done across all the stimuli and as we have already mentioned some of the stimuli were deliberately designed to be non-prototypical, in particular we expected the classroom stimuli where the chairs were arranged in a circle (Figure \ref{fig:classroom-a}) and in a U shaped pattern (Figure \ref{fig:classroom-c} and its horizontally flipped equivalent) to be relatively novel contexts for the subjects. Indeed, if we look at the histogram of the average pairwise $\kappa$ scores by stimuli, Figure \ref{fig:stim-kappa}, it is evident that there is a large variation, ranging from $0$ to $0.8$, in $\kappa$ scores for each stimuli.

\begin{table}[t]
\parbox{.35\linewidth}{
\begin{tabular}{l c c}
% Stim ID & Average Pairwise $\kappa$ & No Region Responses\\
 & Average & No \\
Stim ID & Pairwise $\kappa$ & Reg.\\
\hline
\hline
(h) & 0.5034 & 0\\ % 1_2
(g) & 0.4585 & 1\\ % 1_1
(k) & 0.4548 & 0\\ % 3_3
(f) & 0.3930 & 0\\ % 2_4
(i) & 0.3906 & 0\\ % 3_1
(e-flipped) & 0.3813 & 2\\ % 2_3
(j) & 0.3787 & 1\\ % 3_2
(f-flipped) & 0.3720 & 2\\ % 2_5
(e) & 0.3629 & 1\\ %2_2
(l) & 0.3164& 0\\ % 3_4
(c) & 0.2438 & 2\\ % 2_6
(c-flipped) & 0.2414 & 4\\ %2_7
(a) & 0.0996 & 10 \\ % 2_1
\hline
\end{tabular}
\label{tab:stimuli-kappa-scores}
\caption{The average pairwise $\kappa$ scores and the number of no region response by visual stimuli.}
}
\hfill
\parbox{.65\linewidth}{
\begin{tabular}{c c}
Expected Relationship & Human Results  \\
\hline
\hline
area(g) > area(h) & 291.2 > 276.2 (p<.003) \\
maximumX(f) > maximumX(e) & 176.1 > 167.1   (p<.47) \\
% minimumX(e-flipped) > minimumX(f-flipped) & 336.1 > 321.3 (p<.12) \\
minimumX(e-flipped)  &  \\
 > minimumX(f-flipped) & 336.1 > 321.3 (p<.12) \\
maximumY(i) > maxmimumY(j) & 304.64 > 265.04 (p<.03) \\
maximumY(j) > maxmimumY(k) & 265.04 > 251.71 (p<.47) \\
maximumY(k) = maxmimumY(l) & 265.04 != 290.03 (p<.04) \\
\hline
\end{tabular}
\label{tab:region-location-tests}
\caption{Comparison of the expected relationships to the mean human responses.}
}
\end{table}

@@

Table \ref{tab:stimuli-kappa-scores} lists the average pairwise $\kappa$ score for each of the visual stimuli used in the experiment and the number of no region responses for each stimuli.  It is clear from this table that the less common stimuli have the lowest $\kappa$ scores.  Each of the non-typical regions, the turret +1 arrangement (l), the u-shaped layouts (c and c-flpped), and the circle layout score have kappa scores of less than 0.35. Looking at the classroom results as a whole, these results support the hypothesis that the defintion of CDSRs in a given context is effected by the location and orientation of objects in the environment.  This dependence is observed by the fact that just moving the desks into a circle configuration the CDSR for the front of the room no longer exists.

%Given this relatively large variation across stimuli, which we attribute to the typicality of the depicted context, we computed the average pairwise $\kappa$ scores for each subject when the non-prototypical stimuli trials were excluded. 
%Figure \ref{fig:subj-kappa-proto} presents the histogram of the resulting average pairwise subject  $\kappa$ scores by subject. Not surpringly, with the exclusion on the non-prototypical stimuli the average pairwise $\kappa$ scores increase, with over half the subjects having an average pairwise $\kappa$ score between $0.4$ and $0.6$. $\kappa$ scores in this range can be interpreted as fair to good.

@@

%The analysis of the task 1 trials indicates that, over a range of domains and spatial regions, the responses from the subjects did agree, to a good degree. Furthermore, their results support the hypothesis that the defintion of CDSRs in a given context is effected by the location and orientation of objects in the environment. For example, focusing on the classroom scenes (stimuli, $a$, $c$, $e$, and $f$) the only difference between these trials was the layout of the chairs in the room and yet there is a substantial variation across these stimuli of the agreement between subjects on the specified spatial region, namely \textit{the front of the room}. Situation $f$ resulted in an average pairwise agreement of $0.393$ and in comparision context $a$ had an average pairwise agreement of $0.0996$.

@@

@@

\section{Applicability of Previous Approaches to Cognitive Systems }

Recall from section \ref{} that in addition to being able to represent these regions for a single environment, the cognitive system (1) must recognize regions of the same type in new environments with new contexts, and (2) use the region for multiple reasoning tasks.  While spatial reasoning has been an active topic in AI and robotics for decades, CDSRs present a new problem which requires solutions beyond the current state-of-the-art. Regions are central to QSR work~\cite{Cohn:2001}. Regions can either have crisp boundaries, usually defined by geometric shapes (e.g. the bounding boxes of visually tracked objects~\cite{SridharCohn:10}), or vague boundaries created qualitatively by multiple crisp regions (e.g. the `egg-yolk model'~\cite{Cohn96b}), or by thresholding an underlying spatially-mapped function (e.g. those created by potential fields~\cite{brenneretal07ijcai} or by Gaussian kernels~\cite{burbridge-dearden12}). The fundamental decision when creating a computational model of a CDSR is the method used to represent its boundary. Given their context-dependent nature, boundary definitions based purely on single geometric shapes or a fixed number of points, seem unlikely to model the regions with sufficient accuracy.

However, arbitrary boundaries can be defined using a sequence of geometric primitives (usually straight lines), as in chain codes for image region representations~\cite{Freeman:1961}. This is an approach we followed in previous work \cite{Hawes:2012}, where we defined CDSRs using \textit{anchor points} \cite{DBLP:journals/jetai/KlenkFTK11} as the vertices of the region polygon thereby linking the conceptual regions to perceived objects.   While satisfying the requirements of a cognitive systems solution, applying this model to our results of our experiemnt illustrates two limitations.  First, Regions are not always physically bounded by objects.  In the classroom examples, people frequently did not extend the front of the room all way to the desks.  Second, people consider dynamics part of the context, in the safety example, people did not indicate that it was safe behind the lone turret in front of the others.

The connection between spatial regions and function has been explored in other robotics work, e.g. ~\cite{Karg:2012,Fasola:2013}, but these regions are based on fields anchored at single points in the environment or on a particular object. Therefore, they cannot account for the changes in context shown in this experiment which depend on multiple objects, agents, and simulation.

%Once a particular representation has been modelled in this manner, we were able to find the same type of region in previously unseen rooms using analogical transfer. An example of the representation and transfer approach is Figure \ref{fig:anchor-point-transfer}.  In the know example, a human defines the front of the classroom as a polygon consisting of four anchor points.  Analogical transfer works by aligning structured representations and transferring partially aligned statements from the known example to the new environment.  In this case, the anchor points are transferred from the group of desks in the known example to the corresponding group of desks in the new environment.

%\begin{figure}
%  \includegraphics[width=\columnwidth]{figures/classroom-transfer.png}
%  \caption{The known example defines the front of the room with four anchor points.  By transferring the anchor points to the new environment, the system is able to ground the room's front in its perception.}
%  \label{fig:anchor-point-transfer}
%\end{figure}
